---
title: "Analysis of Institute on Race and Political Economy Guaranteed Income Proposal Marginal Tax Rates"
author: "Jack Landry"
# date: "5/24/2021"
output: html_document
---

<style type="text/css">
body{
  font-family: Unna;
  font-size: 12pt;
}
/* Headers */
h1,h2,h3,h4,h5,h6{
  font-size: 20pt;
  text-align: left;
}

h1.title {
  font-size: 38px;
  text-align: center;
}
h4.author { 
    font-size: 18px;
  text-align: center;
}
h4.date { 
  font-size: 18px;
  text-align: center;
}
</style>



Does not include value of benefits (e.g SNAP, Medicaid) or any state/local taxes. 

```{r, include = FALSE}
theme_jfi <- function () { 
  theme_bw(base_size=12, base_family="Unna") %+replace% 
    theme(panel.grid.minor = element_blank(),
          plot.title = element_text(hjust = 0.5),
          axis.title.y = element_text(angle=0, vjust = 0.5),
          panel.background = element_rect(fill="#fffdfc"), #Not sure if this is actually modifying background
          plot.background = element_rect(fill="#fffdfc"),
          plot.margin=unit(c(.15,.75,.15,.15),"cm")) 
}

library(tidyverse)
library(tidylog)
library(stringr.tools)
library(kableExtra)
library(here)

```

```{r, include=FALSE}
linesize <- 1.25
line_trans <- .5
mt_data <- read_rds(here("inter_data","mt_data_hamilton_for_analysis.Rds"))

mt_data_long_mt_rates <- mt_data %>% 
  select(e00200,couple_status,num_children,num_children_clean,marginal_tax_rate_hamilton_no_eitc,marginal_tax_rate_hamilton_eitc,
         marginal_tax_rate) %>% 
  pivot_longer(cols = c(marginal_tax_rate_hamilton_no_eitc,marginal_tax_rate_hamilton_eitc,marginal_tax_rate),
               names_to = "tax_rate_type",
               values_to = "marginal_tax_rate")

mt_data_long_mt_rates

mt_data_long_mt_rates <- mt_data_long_mt_rates %>% 
  mutate(tax_rate_type=ifelse(tax_rate_type=="marginal_tax_rate", "Status Quo",
                              ifelse(tax_rate_type=="marginal_tax_rate_hamilton_no_eitc","Hamilton Reform\nEIC Repealed", 
                                     ifelse(tax_rate_type=="marginal_tax_rate_hamilton_eitc","Hamilton Reform\nKeep EIC",NA))))

```

```{r, include=FALSE}
singles_marginal_tax_graph <- ggplot(data= mt_data_long_mt_rates %>% filter(couple_status=="Single" & num_children!=6 & e00200!=0), 
       aes(x = e00200, y = marginal_tax_rate, group = tax_rate_type)) +
  geom_line(aes(color = tax_rate_type), size=linesize, alpha = line_trans) +
  facet_wrap(vars(num_children_clean)) +
  labs(y = "Marginal\nTax Rate\nFor Additonal\n$100 Income",
       x = "Labor Income",
       title = "Marginal Tax Rates for Single Filers\n") +
  scale_y_continuous(labels=scales::percent_format(), n.breaks = 5, limits = c(-1,1.3), breaks = seq(-1,1,.5)) +
  scale_x_continuous(labels=scales::dollar_format()) +
  theme_jfi() +
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust=1),
        legend.title = element_blank(),
        # axis.text.y = element_text(angle = 20, vjust = 1, hjust=1),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)))


couples_marginal_tax_graph <- ggplot(data= mt_data_long_mt_rates %>% filter(couple_status=="Couple" & num_children!=6 & e00200!=0), 
       aes(x = e00200, y = marginal_tax_rate, group = tax_rate_type)) +
  geom_line(aes(color=tax_rate_type),size=linesize, alpha = line_trans) +
  facet_wrap(vars(num_children_clean)) +
  labs(y = "Marginal\nTax Rate\nFor Additonal\n$100 Income",
       x = "Labor Income",
       title = "Marginal Tax Rates for Married Filers\n",
       caption = "Jain Family Institute") +
  theme_bw() +
  scale_y_continuous(labels=scales::percent_format(), n.breaks = 5, limits = c(-1,1.3), breaks = seq(-1,1,.5)) +
  scale_x_continuous(labels=scales::dollar_format()) +
  theme_jfi() +
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust=1, size = 8),
        legend.title = element_blank(),
        # axis.text.y = element_text(angle = 20, vjust = 1, hjust=1),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)))

```


```{r, warning = FALSE, message = FALSE, echo=FALSE, figures-marginal-tax-hamilton, fig.show="hold", out.width="50%"}
singles_marginal_tax_graph
couples_marginal_tax_graph
```

```{r, include = FALSE}
linesize <- 1.25
line_trans <- .5



#Want to do the same thing here, include current EITC
mt_data_long_posttax_income <- mt_data %>% 
  select(e00200,couple_status,num_children,num_children_clean,aftertax_hamilton_no_eitc,aftertax_hamilton_eitc,aftertax_income) %>% 
  pivot_longer(cols = c(aftertax_hamilton_no_eitc,aftertax_hamilton_eitc,aftertax_income),
               names_to = "aftertax_income_type",
               values_to = "aftertax_income")


mt_data_long_posttax_income <- mt_data_long_posttax_income %>% 
  mutate(aftertax_income_type=ifelse(aftertax_income_type=="aftertax_income", "Status Quo",
                                     ifelse(aftertax_income_type=="aftertax_hamilton_no_eitc","Hamilton Reform\nEIC Repealed", 
                                            ifelse(aftertax_income_type=="aftertax_hamilton_eitc","Hamilton Reform\nKeep EIC",NA))))




#Need to change the y axis to start at 0
singles_posttax_hamilton_inc_graph <- ggplot(data= mt_data_long_posttax_income %>% filter(couple_status=="Single" & num_children!=6 & e00200!=0 & e00200<=60000), 
       aes(x = e00200, y = aftertax_income, group = aftertax_income_type)) +
  geom_line(aes(color = aftertax_income_type), size=linesize, alpha = line_trans) +
  facet_wrap(vars(num_children_clean)) +
  labs(y = "Aftertax\nIncome",
       x = "Labor Income",
       title = "Pre vs- Post Tax Income for Single Filers\n") +
  scale_y_continuous(labels=scales::dollar_format(), n.breaks = 6, limits = c(0,80000)) +
  scale_x_continuous(labels=scales::dollar_format()) +
  theme_jfi() +
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust=1),
        legend.title = element_blank(),
        # axis.text.y = element_text(angle = 20, vjust = 1, hjust=1),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)))

couples_posttax_hamilton_inc_graph <- ggplot(data= mt_data_long_posttax_income %>% filter(couple_status=="Couple" & num_children!=6 & e00200!=0 & e00200<=80000), 
       aes(x = e00200, y = aftertax_income, group = aftertax_income_type)) +
  geom_line(aes(color = aftertax_income_type), size=linesize, alpha = line_trans) +
  facet_wrap(vars(num_children_clean)) +
  labs(y = "Aftertax\nIncome",
       x = "Labor Income",
       title = "Pre vs- Post Tax Income for Married Filers\n") +
  scale_y_continuous(labels=scales::dollar_format(), n.breaks = 6, limits = c(0,80000)) +
  scale_x_continuous(labels=scales::dollar_format()) +
  theme_jfi() +
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust=1),
        legend.title = element_blank(),
        # axis.text.y = element_text(angle = 20, vjust = 1, hjust=1),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)))

```

```{r, warning = FALSE, message = FALSE, echo=FALSE, figures-posttax-inc-hamilton, fig.show="hold", out.width="50%"}
singles_posttax_hamilton_inc_graph
couples_posttax_hamilton_inc_graph
```

```{r, include=FALSE}

#######################
nber_data <- read_rds(here("inter_data","nber_data_hamilton_for_analysis.Rds"))
#Want to show the difference between 
nber_data_long <- nber_data %>% 
  select(pretax_income,couple_status,num_children,num_children_clean,marginal_tax_rate_hamilton_no_eitc,marginal_tax_rate_hamilton_eitc,
         marginal_tax_rate,marginal_tax_rate_federal_only,marginal_tax_rate_hamilton_no_eitc_fed_only) %>% 
  pivot_longer(cols = c(marginal_tax_rate_hamilton_no_eitc,marginal_tax_rate_hamilton_eitc,marginal_tax_rate,
                        marginal_tax_rate_federal_only,marginal_tax_rate_hamilton_no_eitc_fed_only),
               names_to = "tax_rate_type",
               values_to = "marginal_tax_rate")

#Need to fix this for the new tax rate types
nber_data_long <- nber_data_long %>% 
  mutate(tax_rate_type=ifelse(tax_rate_type=="marginal_tax_rate", "Status Quo",
                              ifelse(tax_rate_type=="marginal_tax_rate_hamilton_no_eitc","Hamilton Reform\nEIC Repealed", 
                                     ifelse(tax_rate_type=="marginal_tax_rate_hamilton_eitc","Hamilton Reform\nKeep EIC",
                                            ifelse(tax_rate_type=="marginal_tax_rate_federal_only","Status Quo Federal Only",
                                                   ifelse(tax_rate_type=="marginal_tax_rate_hamilton_no_eitc_fed_only","Hamilton Reform\nEIC Repealed\nFed Only",NA))))))

#Don' understand why it's not showing a difference

#nber_data_long %>% filter(tax_rate_type=="Status Quo Federal Only"|tax_rate_type=="Status Quo") %>% View()
nber_data_long <- nber_data_long %>% filter(tax_rate_type!="Status Quo Federal Only" & tax_rate_type != "Hamilton Reform\nEIC Repealed\nFed Only")
#How can this be correct????

nber_singles <- ggplot(data= nber_data_long %>% filter(couple_status=="Single" & num_children!=6 & pretax_income!=0 & !is.na(tax_rate_type)), 
       aes(x = pretax_income, y = marginal_tax_rate, group = tax_rate_type)) +
  geom_line(aes(color = tax_rate_type), size=linesize, alpha = line_trans) +
  facet_wrap(vars(num_children_clean)) +
  labs(y = "Marginal\nTax Rate\nFor Additonal\n$100 Income",
       x = "Labor Income",
       title = "Marginal Tax Rates for Single Filers\n") +
  scale_y_continuous(labels=scales::percent_format(), n.breaks = 5, limits = c(-1,1.3), breaks = seq(-1,1,.5)) +
  scale_x_continuous(labels=scales::dollar_format()) +
  theme_jfi() +
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust=1),
        legend.title = element_blank(),
        # axis.text.y = element_text(angle = 20, vjust = 1, hjust=1),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)))

#Issue with single mothers geting EIC
nber_couples <- ggplot(data= nber_data_long %>% filter(couple_status=="Couple" & num_children!=6 & pretax_income!=0 & !is.na(tax_rate_type)), 
       aes(x = pretax_income, y = marginal_tax_rate, group = tax_rate_type)) +
  geom_line(aes(color = tax_rate_type), size=linesize, alpha = line_trans) +
  facet_wrap(vars(num_children_clean)) +
  labs(y = "Marginal\nTax Rate\nFor Additonal\n$100 Income",
       x = "Labor Income",
       title = "Marginal Tax Rates for Married Filers\n") +
  scale_y_continuous(labels=scales::percent_format(), n.breaks = 5, limits = c(-1,1.3), breaks = seq(-1,1,.5)) +
  scale_x_continuous(labels=scales::dollar_format()) +
  theme_jfi() +
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust=1),
        legend.title = element_blank(),
        # axis.text.y = element_text(angle = 20, vjust = 1, hjust=1),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)))

```


```{r, warning = FALSE, message = FALSE, echo=FALSE, figures-posttax-inc-hamilton-nber, fig.show="hold", out.width="50%"}
nber_singles
nber_couples
```


Want to also do pre-post tax income with NBER, add property taxes and rent which are probably important



<!-- Another issues might be incentives to file jointly vs. as single for cohabitators/married... I think for some there'd be an incentive to file seperatly -->
