---
title: "Analysis of Institute on Race and Political Economy Guaranteed Income Proposal"
author: "Jack Landry"
date: "5/24/2021"
output: html_document
---

<style type="text/css">
body{
  font-family: Unna;
  font-size: 12pt;
}
/* Headers */
h1,h2,h3,h4,h5,h6{
  font-size: 20pt;
  text-align: left;
}

h1.title {
  font-size: 38px;
  text-align: center;
}
h4.author { 
    font-size: 18px;
  text-align: center;
}
h4.date { 
  font-size: 18px;
  text-align: center;
}
</style>

## Explination of Program

1. *Single filers:* Benefit of $12,500, phase out starts at $10,000, ends at $50,000
2. *Married/Cohabitating filers:* Benefit of $25,000, phase out starts at $15,000, ends at $70,000
3. *Child benefit:* $4,500/child, no cap, same phase out structure as the main benefit.

```{r, include = FALSE}
#JFI theme for graphics
theme_jfi <- function () { 
  theme_bw(base_size=12, base_family="Unna") %+replace% 
    theme(panel.grid.minor = element_blank(),
          plot.title = element_text(hjust = 0.5),
          axis.title.y = element_text(angle=0, vjust = 0.5),
          panel.background = element_rect(fill="#fffdfc"), #Not sure if this is actually modifying background
          plot.background = element_rect(fill="#fffdfc"),
          plot.margin=unit(c(.15,.75,.15,.15),"cm")) 
}

```

```{r, include=FALSE}
library(tidyverse)
library(RColorBrewer)
reform_eitc_alt <- tibble(eitc = c(12500,12500,0,
                                   17000,17000,0,
                                   21500,21500,0,
                                   26000,26000,0),
                          labor_income = c(0,10000,50000,
                                           0,10000,50000,
                                           0,10000,50000,
                                           0,10000,50000),
                          group = c("0 Children","0 Children","0 Children",
                                    "1 Child", "1 Child", "1 Child",
                                    "2 Children", "2 Children", "2 Children",
                                    "3+ Children","3+ Children","3+ Children")) #Unclear if the phase out threshold is different for children, I don't think it is
reform_eitc_alt <- reform_eitc_alt %>% mutate(`EITC Type`="Guranteed Income EITC",
                                              `Filing Status` = "Single",
                                      group=str_c("Reform EITC: ", group))

reform_eitc_alt_cohabitating <- tibble(eitc = c(25000,25000,0,
                                   29500,29500,0,
                                   34000,34000,0,
                                   38500,38500,0),
                          labor_income = c(0,15000,70000, #Does the start phase out change for cohabitating
                                           0,15000,70000,
                                           0,15000,70000,
                                           0,15000,70000),
                          group = c("0 Children","0 Children","0 Children",
                                    "1 Child", "1 Child", "1 Child",
                                    "2 Children", "2 Children", "2 Children",
                                    "3+ Children","3+ Children","3+ Children")) #Unclear if the phase out threshold is different for children, I don't think it is
reform_eitc_alt_cohabitating <- reform_eitc_alt_cohabitating %>% 
  mutate(`EITC Type`="Guranteed Income EITC",
         `Filing Status` = "Married/Cohabitating",
         group=str_c("Reform EITC Married: ", group))
#Wait it might be a different phase out for the child benefit amount and the regular benefit amount
#The child benefit ends up at the same place, so I don't think so
#For single, child benefit declines at rate of 11.25%, which is 4500
#31.9% for main benefit, which is 12500 


#Have to calculate the new stuff for the reform, maybe I should just make it for the reform I'm actually modeling
original_eitc <- tibble(eitc = c(0,538,538,0,
                                 0,3584,3584,0,
                                 0,5920,5920,0,
                                 0,6660,6660,0),
                        labor_income = c(0,7030,8790,15820,
                                         0,10540,19330,41756,
                                         0,14800,19330,47440,
                                         0,14800,19330,50954),
                        group = c("0 Children","0 Children","0 Children","0 Children",
                                  "1 Child", "1 Child", "1 Child","1 Child",
                                  "2 Children", "2 Children", "2 Children","2 Children",
                                  "3+ Children","3+ Children","3+ Children","3+ Children"))

original_eitc_married <- tibble(eitc = c(0,538,538,0,
                                 0,3584,3584,0,
                                 0,5920,5920,0,
                                 0,6660,6660,0),
                        labor_income = c(0,7030,14680,21710,
                                         0,10540,25220,47646,
                                         0,14800,25220,53330,
                                         0,14800,25220,56844),
                        group = c("0 Children","0 Children","0 Children","0 Children",
                                  "1 Child", "1 Child", "1 Child","1 Child",
                                  "2 Children", "2 Children", "2 Children","2 Children",
                                  "3+ Children","3+ Children","3+ Children","3+ Children"))

original_eitc <- original_eitc %>% mutate(`EITC Type`="Original EITC",
                                          `Filing Status` = "Single",
                                          group=str_c("Original EITC: ", group))

original_eitc_married <- original_eitc_married %>% mutate(`EITC Type`="Original EITC",
                                                          `Filing Status` = "Married/Cohabitating",
                                          group=str_c("Original EITC Married: ", group))






combo_eitc <- bind_rows(original_eitc,original_eitc_married,reform_eitc_alt,reform_eitc_alt_cohabitating)
combo_eitc
hamilton_only <- bind_rows(reform_eitc_alt,reform_eitc_alt_cohabitating)
# hamilton_only <- hamilton_only %>% mutate(group=str_remove(group,"Reform EITC: "),
#                          group=str_remove(group,"Reform EITC Married: "))
#Maybe do singles and married side by side, labeling the lines?
color_for_plots <- "#572dff"


singles <- ggplot(reform_eitc_alt, aes(x = labor_income, y = eitc, group = group)) +
  geom_line(color = color_for_plots) + #Need to control colors so the old is gray
  theme_bw() + 
  labs(y = "Benefit\nSize", x = "\nAdjusted Gross Income", 
       title = "Singles\n") +
  scale_y_continuous(labels=scales::dollar_format(seq(from = 0, to = 50000, by = 10000), accuracy = 1), 
                     breaks = seq(from = 0, to = 50000, by = 10000), limits = c(0,50000)) +
  scale_x_continuous(labels=scales::dollar_format(seq(from = 0, to = 50000, by = 5000), accuracy = 1), 
                     breaks = seq(from = 0, to = 50000, by = 5000), limits = c(0,50000)) +
  theme_jfi() +
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust=1),
        # axis.text.y = element_text(angle = 20, vjust = 1, hjust=1),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)))

cohabitating <- ggplot(reform_eitc_alt_cohabitating, aes(x = labor_income, y = eitc, group = group)) +
  geom_line(color = color_for_plots) + #Need to control colors so the old is gray
  theme_bw() + 
  labs(y = "Benefit\nSize", x = "\nAdjusted Gross Income", 
       title = "Couples\n") +
  scale_y_continuous(labels=scales::dollar_format(seq(from = 0, to = 45000, by = 5000), accuracy = 1), 
                     breaks = seq(from = 0, to = 45000, by = 5000), limits = c(0,45000)) +
  scale_x_continuous(labels=scales::dollar_format(seq(from = 0, to = 70000, by = 15000), accuracy = 1), 
                     breaks = seq(from = 0, to = 70000, by = 15000), limits = c(0,70000)) +
  theme_jfi() +
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust=1),
        # axis.text.y = element_text(angle = 20, vjust = 1, hjust=1),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)))

cohabitating

size_for_annotation <- 2.5
singles <- singles + 
  annotate("label", label = "0 Children", x = 5000, y = 14000, size = size_for_annotation, colour = color_for_plots,
           family= "Unna") +
  annotate("label", label = "1 Child", x = 5000, y = 18500, size = size_for_annotation, colour = color_for_plots,
           family= "Unna") +
  annotate("label", label = "2 Children", x = 5000, y = 23000, size = size_for_annotation, colour = color_for_plots,
           family= "Unna") +
  annotate("label", label = "3 Children", x = 5000, y = 27500, size = size_for_annotation, colour = color_for_plots,
           family= "Unna") 
singles


cohabitating <- cohabitating + 
  annotate("label", label = "0 Children", x = 5000, y = 27000, size = size_for_annotation, colour = color_for_plots,
           family= "Unna") +
  annotate("label", label = "1 Child", x = 5000, y = 31500, size = size_for_annotation, colour = color_for_plots,
           family= "Unna") +
  annotate("label", label = "2 Children", x = 5000, y = 36000, size = size_for_annotation, colour = color_for_plots,
           family= "Unna") +
  annotate("label", label = "3 Children", x = 5000, y = 40500, size = size_for_annotation, colour = color_for_plots,
           family= "Unna") 
cohabitating

library(cowplot)

library(patchwork)
hamilton_by_kids <- (singles + cohabitating) +
  plot_annotation(title = 'Guaranteed Income Benefit Size', theme = theme(plot.title = element_text(hjust = 0.5, size = 16), text=element_text(family="Unna")))



#Need to get the legend going nicely
guranteed_income_plot <- ggplot(combo_eitc, aes(x = labor_income, y = eitc, group = group, color = `EITC Type`, linetype=`Filing Status`)) +
  geom_line() + #Need to control colors so the old is gray
  scale_color_manual(values=c("#572dff","gray")) +
  theme_bw() + 
  labs(y = "Benefit\nSize", x = "\nAdjusted Gross Income", 
       title = "Guranteed Income Benefit\n Vs. Standard EITC",
       caption = "Jain Family Institute") +
  scale_y_continuous(labels=scales::dollar_format(), n.breaks = 9) +
  scale_x_continuous(labels=scales::dollar_format()) +
  theme_jfi()
```

```{r, echo=FALSE, fig.align="center", dpi = 300}
hamilton_by_kids
```

The Hamilton proposal differs from the standard EITC in several ways. The most important change in the lack of phase in--Hamilton gives people with no taxable earnings the full benefit, while the standard EITC would give people nothing. Hamilton's program is also more generous, almost everyone will receive a larger benefit than the existing EITC. However, the benefits have a much more aggressive phase out relative to the standard EITC. The phase out are especially large families with more children--these families see a larger benefit but see the same phase in and phase out thresholds as people with no children, which necessitates a much steeper slope.

```{r, echo=FALSE, fig.align="center", out.width = "80%", dpi = 300}
guranteed_income_plot
```


## Explaining Hamilton Analysis Errors

I estimate that if everyone eligible for the program received the full benefit they were entitled to, the program would cost \$1.4 trillion dollars.^[This uses data on incomes from 2019, the same year of analysis of the Hamilton program.] Hamilton estimates the program cost at \$876 billion dollars. This discrepancy is caused by several different facets of modeling assumptions, which I detail in the following sections. 

```{r, include = FALSE}
median_income_change_hamilton <- 64742-59051
median_income_change_hamilton
```

Although I estimate the Hamilton program is significantly more expensive than advertised, that does not necessarily mean that they underestimate the cost of eliminating poverty. The larger cost I estimate means that more benefits are going out the door. For instance, Hamilton estimates that the program will boost median family income by \$5,691, while my estimate is \$11,881.^[My estimate is of median family income measure on the individual level, e.g. weighted by family size. Need to check if Hamilton clarifies that.] If you change the program rules, it may be possible achieve the same poverty reductions with costs close to what Hamilton estimates.^[In practice, it might be impossible to achieve the same poverty impacts conditional on size due to difficulties targeting poor families with a tax code that emphasizes individual income.]

### Phase out rates

```{r, echo=FALSE}
#Maybe try to make a table that reproduces their table 1
```

Hamilton writes that the phase out rate for singles is 31.9% and the phase out rate for couples is 46.4%. These proposed phase out rates are slightly inconsistent with how they describe the program rules. Take singles: with a benefit of \$12,500, a phase out that starts at \$10,000 and ends at \$50,000, the phase out rate should be 31.25%. This comes from the formula for calculating a slope. $$m 
	= \frac{\Delta y}{\Delta x} 
	= \frac{y_2 - y_1}{x_2 - x_1}
	= \frac{12,500 - 0}{10,000 - 50,000}
	= 31.25\%$$ 
Similarly, the phase out for couples is $$m = \frac{\Delta y}{\Delta x} 
	= \frac{y_2 - y_1}{x_2 - x_1}
	= \frac{25,000 - 0}{15,000 - 70,000}
	= 45.45\%$$ 
The general benefit calculation is for people with a income between the phase out start and end points is $$benefit = max\_benefit - (income - phase\_out\_start) * phase\_out\_rate$$
where the phase out rate comes from the calculation for the slope$$m 
	= \frac{\Delta y}{\Delta x} 
	= \frac{y_2 - y_1}{x_2 - x_1}
	= \frac{max\_benefit - 0}{phase\_out\_start - phase\_out\_end}
	= phase\_out\_rate$$
The maximum benefit is either \$12,500 or \$25,000, depending on filing status, with the addition of $4,500 times the number of children in the household.

These errors in calculating the phase out rate lead to small discrepancies in the benefit calculation of hypothetical households with $40,000 of income, shown in Hamilton's Table 1. Note that the benefit values in the table are also inconsistent with Hamilton's phase outs, but none of the errors are larger than \$500.

<!-- Should I add phase out's implied by their phase out percentages? -->


```{r, include = FALSE}
library(tidyverse)
library(tidylog)
library(stringr.tools)
library(kableExtra)
library(here)

hamilton_singles <- read_rds(here("inter_data","singles_hamilton.rds"))
main_benefit <- hamilton_singles$hamilton_money[1]
hamilton_singles <- hamilton_singles %>% 
  mutate(hamilton_money=ifelse(hamilton_money!=main_benefit,hamilton_money-main_benefit,hamilton_money)) %>% 
  mutate(`Corrected Refund Amount`=hamilton_money) %>% 
  select(`Corrected Refund Amount`)

hamilton_table <- tibble(`Refund Recepient` = c("Adult", "Child 1", "Child 2"),
                         `Refund Amount Hamilton` = c(3269,1159,2318))
hamilton_table <- bind_cols(hamilton_table,hamilton_singles) %>% mutate(Difference=`Refund Amount Hamilton`-`Corrected Refund Amount`)
hamilton_table

hamilton_table <- hamilton_table %>% mutate(across(where(is.numeric), round, 0)) %>% 
  mutate(across(where(is.numeric),as.character)) %>% #What about comma insertion
  mutate(across(`Refund Amount Hamilton`:Difference, ~prettyNum(.x,big.mark=",")))  %>% 
  mutate(across(`Refund Amount Hamilton`:Difference, ~str_prefix(.x,"$"))) 


#Note that they both get to claim the kid...
hamilton_table_couples <- tibble(`Refund Recepient` = c("Couple", "Child 1", "Child 2"),
                                 `Refund Amount Hamilton` = c(14048,2475,4958))

hamilton_couples <- read_rds(here("inter_data","couples_hamilton.rds"))
main_benefit_couples <- hamilton_couples$hamilton_money[1]
main_benefit <- hamilton_singles$hamilton_money[1]
hamilton_couples <- hamilton_couples %>% 
  mutate(hamilton_money=ifelse(hamilton_money!=main_benefit_couples,hamilton_money-main_benefit_couples,hamilton_money)) %>% 
  mutate(`Corrected Refund Amount`=hamilton_money) %>% 
  select(`Corrected Refund Amount`)

hamilton_table_couples <- bind_cols(hamilton_table_couples,hamilton_couples) %>% mutate(Difference=`Refund Amount Hamilton`-`Corrected Refund Amount`)
hamilton_table_couples


hamilton_table_couples <- hamilton_table_couples %>% mutate(across(where(is.numeric), round, 0)) %>% 
  mutate(across(where(is.numeric),as.character)) %>% #What about comma insertion
  mutate(across(`Refund Amount Hamilton`:Difference, ~prettyNum(.x,big.mark=",")))  %>% 
  mutate(across(`Refund Amount Hamilton`:Difference, ~str_prefix(.x,"$"))) 

```

```{r, echo=FALSE}
knitr::kable(hamilton_table, format = "html", 
             caption = "<center><strong>Singles</strong></center>",
             escape = FALSE) %>% 
  kable_styling(html_font = "Unna",
                full_width = FALSE,
                position = "center")
```

```{r, echo=FALSE}
knitr::kable(hamilton_table_couples, format = "html", 
             caption = "<center><strong>Couples</strong></center>",
             escape = FALSE) %>% 
  kable_styling(html_font = "Unna",
                full_width = FALSE,
                position = "center")
```

### Household Construction

To model benefits and estimate cost, Hamilton uses the Survey of Consumer Finances (SCF) summary extract data. This is a commonly used, high-quality survey with detailed information on income and wealth. However, the summary extract have data on households, not individuals. For tax filing purposes (and how the Hamilton program is described), the only income data relevant is your personal income or your combined income with a cohabitating partner. When there are multiple adult individuals in a household, they will often file taxes separately based on their individual incomes, but Hamilton only has information on household combined income. Since combined incomes are higher than individual incomes by construction, using household income to determine eligibility decreases the number of people who appear eligible and thus the estimated cost of the program. In the appendix, Hamilton describes a correction for households with three or more adults:

> Finally, the SCF data also accounts for households with three or more adults, while the tax system only allows for single and two-adult filings. To estimate the economic effects of our program for households in the SCF data with three or more adults, we adopt a simplifying convention: households with three or more adults receive up to \$30,000 total for those households with income up to \$30,000, declining at a rate of 50 percent above \$30,000 to fully phase out at $80,000 in total household income across all adults. Because this category only represents a small proportion of total households, the choice of how to estimate the grant amounts for these households will not significantly affect the overall program cost. Under an alternative assumption that every non-married adult (with two or more adults in the household) file separately, that divides household income in a way that maximizes the refund value, our guaranteed income program cost still remains under \$1 trillion.

Add details about how this matters in practice with numbers.

<!-- There is info on number of adults in the non-summary extract data -->

<!-- I think this whole thing speaks to the difficulties of targeting only poor families without overhauling the tax code that mostly emphasizes individuals. There are a bunch of adults who have little taxable income in families that are not poor who receive income under this plan. Hamilton's family income based targeting doesn't give money to these people, but my individual level income data does.  -->


### Adjusted Gross Income vs. Total Income

To determine eligibility for the program, the Hamilton plan calls for using adjusted gross income (a.k.a taxable income). However, in their analysis of the SCF, they use a total income measure, which does not adjust for types of income that are not taxable. By construction, this income measure will increase people's measure income compare to their taxable income an underestimate eligability for the program.

Add details using CPS

## Analysis of Program

### Poverty Impacts 

### Average Tax Rate Analysis

```{r, include = FALSE}
mt_data <- read_rds(here("inter_data","mt_data_hamilton_for_analysis.Rds"))



#Average tax rate singles
singles_avg_hamilton <- ggplot(data= mt_data %>% filter(couple_status=="Single" & num_children!=6 & e00200>=5000), 
       aes(x = e00200, y = tax_rate_hamilton_no_eitc, group = num_children_clean)) +
  geom_line(aes(color=num_children_clean)) +
  labs(y = "Average Tax Rate",
       x = "Labor Income",
       title = "Average Tax Rates for Single Filers\nHamilton Guaranteed Income Progam\n",
       caption = "Jain Family Institute") +
  # scale_color_discrete(name = "Number of Children") +
  scale_color_brewer(palette = "Set1", name = "Number of Children") +
  scale_y_continuous(labels=scales::percent_format(), n.breaks = 8) +
  scale_x_continuous(labels=scales::dollar_format(seq(from = 5000, to = 100000, by = 10000)),
                     breaks=seq(from = 5000, to = 100000, by = 10000)) +
  theme_jfi() +
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust=1),
        # axis.text.y = element_text(angle = 20, vjust = 1, hjust=1),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)))


#Average tax rate coupes
couples_avg_hamilton <- ggplot(data= mt_data %>% filter(couple_status=="Couple" & num_children!=6 & e00200>=5000), 
       aes(x = e00200, y = tax_rate_hamilton_no_eitc, group = num_children_clean)) +
  geom_line(aes(color=num_children_clean)) +
  labs(y = "Average Tax Rate",
       x = "Labor Income",
       title = "Average Tax Rates for Married Filers\nHamilton Guaranteed Income Progam\n",
       caption = "Jain Family Institute") +
  # scale_color_discrete(name = "Number of Children") +
  scale_color_brewer(palette = "Set1", name = "Number of Children") +
  scale_y_continuous(labels=scales::percent_format(), n.breaks = 8) +
  scale_x_continuous(labels=scales::dollar_format(seq(from = 5000, to = 100000, by = 10000)),
                     breaks=seq(from = 5000, to = 100000, by = 10000)) +
  theme_jfi() +
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust=1),
        # axis.text.y = element_text(angle = 20, vjust = 1, hjust=1),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)))
```

```{r, warning = FALSE, message = FALSE, echo=FALSE, figures-avg-tax-hamilton, fig.show="hold", out.width="50%"}
couples_avg_hamilton
singles_avg_hamilton
```


### Marginal Tax Rate Analysis

I should also plot pre tax vs. post tax income. 

```{r, include=FALSE}
linesize <- 1.25
line_trans <- .5

singles_marginal_tax_graph <- ggplot(data= mt_data %>% filter(couple_status=="Single" & num_children!=6), 
       aes(x = e00200, y = marginal_tax_rate_hamilton_eitc)) +
  geom_line(size=linesize, alpha = line_trans) +
  facet_wrap(vars(num_children_clean)) +
  labs(y = "Marginal\nTax Rate",
       x = "Labor Income",
       title = "Marginal Tax Rates for Single Filers\n") +
  scale_y_continuous(labels=scales::percent_format(), n.breaks = 6) +
  scale_x_continuous(labels=scales::dollar_format()) +
  theme_jfi() +
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust=1),
        # axis.text.y = element_text(angle = 20, vjust = 1, hjust=1),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)))
# ggsave(here("new_code","analysis","hamilton-analysis","plots","marginal_tax_rate_hamilton_no_eitc_singles.png"),
#        width = 4, height = 2, units = "cm", dpi = 1000)


couples_marginal_tax_graph <- ggplot(data= mt_data %>% filter(couple_status=="Couple" & num_children!=6), 
       aes(x = e00200, y = marginal_tax_rate_hamilton_no_eitc)) +
  geom_line(size=linesize, alpha = line_trans) +
  facet_wrap(vars(num_children_clean)) +
  labs(y = "Marginal\nTax Rate\nFor Additonal\n$100 Income",
       x = "Labor Income",
       title = "Marginal Tax Rates for Married Filers\n",
       caption = "Jain Family Institute") +
  theme_bw() +
  scale_y_continuous(labels=scales::percent_format(), n.breaks = 6) +
  scale_x_continuous(labels=scales::dollar_format()) +
  theme_jfi() +
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust=1),
        # axis.text.y = element_text(angle = 20, vjust = 1, hjust=1),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)))

```


```{r, warning = FALSE, message = FALSE, echo=FALSE, figures-marginal-tax-hamilton, fig.show="hold", out.width="50%"}
singles_marginal_tax_graph
couples_marginal_tax_graph
```

```{r, include = FALSE}
linesize <- 1.25
line_trans <- .5
singles_posttax_hamilton_inc_graph <- ggplot(data= mt_data %>% filter(couple_status=="Single" & num_children!=6), 
       aes(x = e00200, y = aftertax_hamilton_no_eitc)) +
  geom_line(size=linesize, alpha = line_trans) +
  facet_wrap(vars(num_children_clean)) +
  labs(y = "Aftertax Income\nWith Hamilton Benefit",
       x = "Labor Income",
       title = "Pre vs- Post Tax Income for Single Filers\n") +
  scale_y_continuous(labels=scales::dollar_format(), n.breaks = 6) +
  scale_x_continuous(labels=scales::dollar_format()) +
  theme_jfi() +
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust=1),
        # axis.text.y = element_text(angle = 20, vjust = 1, hjust=1),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)))

couples_posttax_hamilton_inc_graph <- ggplot(data= mt_data %>% filter(couple_status=="Couple" & num_children!=6), 
       aes(x = e00200, y = aftertax_hamilton_no_eitc)) +
  geom_line(size=linesize, alpha = line_trans) +
  facet_wrap(vars(num_children_clean)) +
  labs(y = "Aftertax Income\nWith Hamilton Benefit",
       x = "Labor Income",
       title = "Pre vs- Post Tax Income for Married Filers\n") +
  scale_y_continuous(labels=scales::dollar_format(), n.breaks = 6) +
  scale_x_continuous(labels=scales::dollar_format()) +
  theme_jfi() +
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust=1),
        # axis.text.y = element_text(angle = 20, vjust = 1, hjust=1),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)))

```

```{r, warning = FALSE, message = FALSE, echo=FALSE, figures-posttax-inc-hamilton, fig.show="hold", out.width="50%"}
singles_posttax_hamilton_inc_graph
couples_posttax_hamilton_inc_graph
```
