---
title: "50% EITC Replication"
author: "Jack Landry"
date: "5/26/2021"
output: html_document
---

<style type="text/css">
body{
  font-family: Unna;
  font-size: 12pt;
}
/* Headers */
h1,h2,h3,h4,h5,h6{
  font-size: 20pt;
  text-align: center;
}

h1.title {
  font-size: 38px;
  text-align: center;
}
h4.author { 
    font-size: 18px;
  text-align: center;
}
h4.date { 
  font-size: 18px;
  text-align: center;
}
</style>

```{r setup, include=FALSE, cashe=TRUE}
library(tidyverse)
library(ipumsr)
library(tidylog)
library(kableExtra)
library(stringr.tools)
library(here)
here()
#Tax units sent to calculator
ids <- read_csv(here("outputs","ids.csv")) 
#New policy results file
new_policy <- read_csv(here("tax-calculator-outputs","earnie-reform.csv"))
#Status quo policy results file
status_quo_poicy <- read_csv(here("tax-calculator-code","c17_taxsim-17-#-#-#.csv"))

new_policy <- new_policy %>% select(RECID,aftertax_income) %>% rename(new_pol_aftertax_income=aftertax_income)
status_quo_poicy <- status_quo_poicy %>% select(RECID,aftertax_income) %>%
  rename(stats_quo_pol_aftertax_income=aftertax_income)

combo_data <- ids %>% left_join(status_quo_poicy) %>% left_join(new_policy)
combo_data <- combo_data %>% mutate(diff_aftertax_income=new_pol_aftertax_income - stats_quo_pol_aftertax_income)

#Now bring in CPS ASEC data raw data for analysis
ddi <- read_ipums_ddi(here("cps_data","cps_00042.xml"))
ipum <-  read_ipums_micro(ddi)
names(ipum) <- tolower(names(ipum))
#ipum <- ipum %>% select(year,age,serial,pernum,spmwt,spmthresh,spmtotres,spmfamunit) #Change this in the future

#"Not every ASEC record became a tax unit, so make sure you keep master non-matches" (From Earnie)
#So ASEC records that are not tax units had no change in income
#Also, why didn't they become tax units, how is nonfiling handled?
ipum <- ipum  %>% left_join(combo_data, by = c("serial","pernum"))

```

```{r, include=FALSE}

ipum <- ipum %>% mutate(no_aftertax_income=ifelse(is.na(diff_aftertax_income),1,0),
                        diff_aftertax_income=ifelse(is.na(diff_aftertax_income),0,diff_aftertax_income)) 
#No match in tax data, no change in income, so setting missings to zero, which is how this egen sum thing is handeled in stata. Do retain an indicator for this. Need to understand what's going on in the tax data better to know why half of observations don't match
ipum <- ipum %>% group_by(spmfamunit) %>% 
  mutate(diff_aftertax_income_family=sum(diff_aftertax_income)) %>% 
  ungroup()
# 
# sum(is.na(ipum$diff_aftertax_income)) #Ton of missings
# #New, setting difference in aftertax income to 0 if not a filing unit
# ipum <-ipum %>% 
#   mutate(diff_aftertax_income_family=ifelse(is.na(diff_aftertax_income_family),0,diff_aftertax_income_family))

#Need to look at how this is done again, look at what these variables are
#SPMTOTRES records the total economic resources of families and is used to construct the Supplemental Poverty Measure (SPM)
#SPMTHRESH records the poverty threshold used to calculate the Supplemental Poverty Measure (SPM) for a family.
ipum <- ipum %>% mutate(spm1 = spmtotres < spmthresh,
                spm2 = (spmtotres + diff_aftertax_income_family) < spmthresh,
                out_of_spm = ifelse((spm2 == 0 & spm1 == 1),1,0),
                out_of_spm = ifelse(is.na(out_of_spm),0,out_of_spm)) #I think this is right but not sure 
```


```{r, include=FALSE}
#Need to get the total HH numbers too
#Need to round this whole thing, improve names, etc.
tables_test <- ipum %>% summarise(
  spm1_mean = weighted.mean(spm1, spmwt),
  spm2_mean = weighted.mean(spm2, spmwt),
  out_of_spm_mean = weighted.mean(out_of_spm, spmwt))
#Trying to do this for children, poverty is at the hh level, but data is at the individual level
#So for children, measuring the reduction in the % of children whose HH are in poverty
child_poverty <- ipum %>% filter(age<18) %>% summarise(
  spm1_mean = weighted.mean(spm1, spmwt),
  spm2_mean = weighted.mean(spm2, spmwt),
  out_of_spm_mean = weighted.mean(out_of_spm, spmwt)) 
percent_poverty_reduction <- (tables_test$spm1_mean-tables_test$spm2_mean)/tables_test$spm2_mean
percent_poverty_reduction <- round(percent_poverty_reduction, 3)*100
percentage_point_poverty_reduction <- round(tables_test$out_of_spm_mean, 3)*100
tables_test <- tables_test %>% rename(`Status Quo Poverty` = spm1_mean,
                       `EITC Reform Poverty` = spm2_mean,
                       `Poverty Reduction` = out_of_spm_mean
                       )


tables_test <- tables_test %>% 
  mutate(across(where(is.numeric), round, 3)) %>% 
  mutate(across(where(is.numeric), ~ .x*100)) %>% 
  mutate(across(where(is.numeric),as.character)) %>% 
  mutate(across(where(is.character), ~str_suffix(.x,"%")))
  

```
<!-- Calculating changes in poverty, want to make these nice tables -->
<!-- Want to change the html to mimic Jain style as much as possible -->
<!-- Look at exsample here: https://www.jainfamilyinstitute.org/news/robust-evidence-for-1400-relief-and-recovery-checks/ -->
## Poverty Impacts of Expanded EITC

Calculating changes in poverty from random changes to the EITC I made up. Want to add a column for catagory, and then add rows for overall, children, and more groups. Note that poverty measure is at the HH level, and most catagories would be at the individual level.

```{r, echo=FALSE}
tables_test %>%
  kbl() %>%
  kable_minimal() %>% 
  kable_styling(html_font = "Unna",
                full_width = FALSE,
                position = "center")

```



```{r, include=FALSE}
#Getting number of HHs in poverty or whatever
#https://cran.r-project.org/web/packages/ipumsr/vignettes/ipums-cps.html
hhs_poverty_status_quo <- ipum %>% 
  group_by(spm1) %>%
  summarize(hhs_in_poverty = sum(spmwt)) %>% 
  filter(spm1==TRUE) %>% pull()
hhs_poverty_reform <- ipum %>% 
  group_by(spm2) %>%
  summarize(hhs_in_poverty = sum(spmwt)) %>% 
  filter(spm2==TRUE) %>% pull()

num_of_hh_out_of_poverty <- hhs_poverty_status_quo-hhs_poverty_reform
num_of_hh_out_of_poverty <- prettyNum(num_of_hh_out_of_poverty,big.mark=",",scientific=FALSE) #Making this display nicely
```
This reform caused a `r num_of_hh_out_of_poverty[1]` household poverty reduction, a `r percent_poverty_reduction[1]`% decrease (a `r percentage_point_poverty_reduction[1]` percentage point reduction). 

