---
title: "Tax Changes Template"
author: "Jack Landry"
date: "5/24/2021"
output: html_document
---

<style type="text/css">
body{
  font-family: Unna;
  font-size: 12pt;
}
/* Headers */
h1,h2,h3,h4,h5,h6{
  font-size: 20pt;
  text-align: center;
}

h1.title {
  font-size: 38px;
  text-align: center;
}
h4.author { 
    font-size: 18px;
  text-align: center;
}
h4.date { 
  font-size: 18px;
  text-align: center;
}
</style>

```{r setup, include=FALSE, cashe=TRUE}
#Not sure i
library(tidyverse)
library(ipumsr)
library(tidylog)
library(kableExtra)
library(stringr.tools)
library(here)
here()
#Tax units sent to calculator
ids <- read_csv(here("outputs","ids.csv")) 
#New policy results file
new_policy <- read_csv(here("tax-calculator-outputs","eitc_reform.csv"))
#Status quo policy results file
status_quo_poicy <- read_csv(here("tax-calculator-code","c17_taxsim-17-#-#-#.csv"))

#Need to change this to keep all these output variabes, but add a prefix or something
names(new_policy) <- paste("reform", names(new_policy), sep = "_")

# new_policy <- new_policy %>% rename(new_pol_aftertax_income=raftertax_income)
names(status_quo_poicy) <- paste("status_quo", names(status_quo_poicy), sep = "_")
# status_quo_poicy <- status_quo_poicy %>%
#   rename(stats_quo_pol_aftertax_income=aftertax_income)

combo_data <- ids %>% left_join(status_quo_poicy, by = c("RECID" = "status_quo_RECID")) %>% 
  left_join(new_policy, by = c("RECID" = "reform_RECID"))
combo_data <- combo_data %>% mutate(diff_aftertax_income=reform_aftertax_income - status_quo_aftertax_income)

#Now bring in CPS ASEC data raw data for analysis
ddi <- read_ipums_ddi(here("cps_data","cps_00042.xml"))
ipum <-  read_ipums_micro(ddi)
names(ipum) <- tolower(names(ipum))
#ipum <- ipum %>% select(year,age,serial,pernum,spmwt,spmthresh,spmtotres,spmfamunit) #Change this in the future

#"Not every ASEC record became a tax unit, so make sure you keep master non-matches" (From Earnie)
#So ASEC records that are not tax units had no change in income
#Also, why didn't they become tax units, how is nonfiling handled?
ipum <- ipum  %>% left_join(combo_data, by = c("serial","pernum"))
```

```{r, include=FALSE}
ipum <- ipum %>% mutate(no_aftertax_income=ifelse(is.na(diff_aftertax_income),1,0),
                        diff_aftertax_income=ifelse(is.na(diff_aftertax_income),0,diff_aftertax_income)) 

#Need to be careful about how missings are treated, I think if they aren't in the tax data the change should be 0
ipum <- ipum %>% group_by(spmfamunit) %>% 
  mutate(diff_aftertax_income_family=sum(diff_aftertax_income)) %>% 
  ungroup()

#SPMTOTRES records the total economic resources of families and is used to construct the Supplemental Poverty Measure (SPM)
#SPMTHRESH records the poverty threshold used to calculate the Supplemental Poverty Measure (SPM) for a family.

ipum <- ipum %>% mutate(deep_poverty=spmthresh*.5,
                baseline_poverty = spmtotres < spmthresh,
                post_reform_poverty = (spmtotres + diff_aftertax_income_family) < spmthresh,
                out_of_poverty = ifelse((post_reform_poverty == 0 & baseline_poverty == 1),1,0),
                out_of_poverty = ifelse(is.na(out_of_poverty),0,out_of_poverty),
                baseline_deep_poverty=spmtotres < deep_poverty,
                post_reform_deep_poverty=(spmtotres + diff_aftertax_income_family) < deep_poverty,
                out_of_deep_poverty = ifelse((post_reform_deep_poverty == 0 & baseline_deep_poverty == 1),1,0),
                out_of_poverty = ifelse(is.na(out_of_deep_poverty),0,out_of_deep_poverty)) 



```

```{r, include = FALSE}
#Making a table out of my EITC changes
eitc <- tibble(eitc = c(5000,5000,0,
                        10000,10000,0,
                        15000,15000,0,
                        20000,20000,0), 
               labor_income = c(0,30000,95359,
                                0,30000,92578.2,
                                0,30000,101025,
                                0,30000,124967), 
               group = c("0 Children","0 Children","0 Children",
                         "1 Child", "1 Child", "1 Child",
                         "2 Children", "2 Children", "2 Children",
                         "3+ Children","3+ Children","3+ Children"))

library(ggrepel)
eitc_plot <- ggplot(eitc, aes(x = labor_income, y = eitc, group = group)) +
  geom_line(color = "#572dff") +
  theme_bw() + 
  labs(y = "EITC  \nBenefit  ", x = "\nLabor Income", 
       title = "New EITC Benefit For Single Filers",
       caption = "Jain Family Institute") +
  scale_y_continuous(labels=scales::dollar_format()) +
  scale_x_continuous(labels=scales::dollar_format()) +
  theme(plot.title = element_text(hjust = 0.5),
        text=element_text(family="Unna"), #Want to change this to Jain font, need to ask for it
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(angle=0, vjust = 0.5),
        panel.background = element_rect(fill="#fffdfc"), #Not sure if this is actually modifying background
        plot.background = element_rect(fill="#fffdfc"),
        plot.margin=unit(c(.15,.75,.15,.15),"cm")) 

eitc_plot <- eitc_plot + 
  annotate("label", label = "0 Children", x = 15000, y = 6000, size = 3.5, colour = "#572dff",
           family= "Unna") +
  annotate("label", label = "1 Child", x = 15000, y = 11000, size = 3.5, colour = "#572dff",
           family= "Unna") +
  annotate("label", label = "2 Children", x = 15000, y = 16000, size = 3.5, colour = "#572dff",
           family= "Unna") +
  annotate("label", label = "3+ Children", x = 15000, y = 21000, size = 3.5, colour = "#572dff",
           family= "Unna") 
```

## Explanation of Reform

What does the reform look like? Here, I pick pretty arbitrary numbers to remove the phase in, increase the max benefit, and increase the phase out threshold. I didn't touch the phase out rate (double check), should I do that in the future? I think I want to add lines for married filers, as well as a comparison to the status-quo EITC. Maybe add some light gray lines for the status-quo? Labels might get messy.
```{r, echo=FALSE, fig.align="center"}
eitc_plot
```

Should maybe add how post-tax income varies with labor income pre-post reform by number of kids.

```{r, include=FALSE}
#Add deep poverty to this
overall_poverty <- ipum %>% summarise(
  baseline_poverty_mean = weighted.mean(baseline_poverty, spmwt),
  reform_poverty_mean = weighted.mean(post_reform_poverty, spmwt),
  reform_poverty_reduction = weighted.mean(out_of_poverty, spmwt))
#Trying to do this for children, poverty is at the hh level, but data is at the individual level
#So for children, measuring the reduction in the % of children whose HH are in poverty
child_poverty <- ipum %>% filter(age<18) %>% summarise(
  baseline_poverty_mean = weighted.mean(baseline_poverty, spmwt),
  reform_poverty_mean = weighted.mean(post_reform_poverty, spmwt),
  reform_poverty_reduction = weighted.mean(out_of_poverty, spmwt)) 

#Separate table for deep poverty, if I do any more thresholds make this a function
overall_deep_poverty <- ipum %>% summarise(
  baseline_deep_poverty_mean = weighted.mean(baseline_deep_poverty,spmwt),
  reform_deep_poverty_mean = weighted.mean(post_reform_deep_poverty,spmwt),
  reform_deep_poverty_reduction = weighted.mean(out_of_deep_poverty,spmwt)) %>% 
  mutate(Group = "Overall")

child_deep_poverty <- ipum %>% filter(age<18) %>% summarise(
  baseline_deep_poverty_mean = weighted.mean(baseline_deep_poverty,spmwt),
  reform_deep_poverty_mean = weighted.mean(post_reform_deep_poverty,spmwt),
  reform_deep_poverty_reduction = weighted.mean(out_of_deep_poverty,spmwt)) %>% 
  mutate(Group = "Children")

#Want to do more groups here


percent_poverty_reduction <- (overall_poverty$baseline_poverty_mean-overall_poverty$reform_poverty_mean)/overall_poverty$baseline_poverty_mean

percent_poverty_reduction <- round(percent_poverty_reduction, 3)*100
percentage_point_poverty_reduction <- round(overall_poverty$reform_poverty_reduction, 3)*100
child_poverty <- child_poverty %>% mutate(Group="Children")
overall_poverty <- overall_poverty %>% mutate(Group="Overall")
poverty_table <- bind_rows(overall_poverty,child_poverty)
poverty_table <- poverty_table %>% rename(`Status Quo Poverty` = baseline_poverty_mean,
                       `EITC Reform Poverty` = reform_poverty_mean,
                       `Poverty Reduction` = reform_poverty_reduction
                       )


poverty_table <- poverty_table %>% 
  mutate(across(where(is.numeric), round, 3)) %>% 
  mutate(across(where(is.numeric), ~ .x*100)) %>% 
  mutate(across(where(is.numeric),as.character)) %>% 
  mutate(across(`Status Quo Poverty`:`Poverty Reduction`, ~str_suffix(.x,"%"))) %>% 
  relocate(Group)

deep_poverty_table <- bind_rows(overall_deep_poverty,child_deep_poverty)
deep_poverty_table <- deep_poverty_table %>% rename(`Status Quo Deep Poverty` = baseline_deep_poverty_mean,
                       `EITC Reform Deep Poverty` = reform_deep_poverty_mean,
                       `Deep Poverty Reduction` = reform_deep_poverty_reduction
                       )

deep_poverty_table <- deep_poverty_table %>% 
  mutate(across(where(is.numeric), round, 3)) %>% 
  mutate(across(where(is.numeric), ~ .x*100)) %>% 
  mutate(across(where(is.numeric),as.character)) %>% 
  mutate(across(`Status Quo Deep Poverty`:`Deep Poverty Reduction`, ~str_suffix(.x,"%"))) %>% 
  relocate(Group)


```
<!-- Calculating changes in poverty, want to make these nice tables -->
<!-- Want to change the html to mimic Jain style as much as possible -->
<!-- Look at example here: https://www.jainfamilyinstitute.org/news/robust-evidence-for-1400-relief-and-recovery-checks/ -->
## Poverty Impacts of Expanded EITC

Calculating changes in poverty from random changes to the EITC I made up. Note that poverty measure is at the HH level, and most categories would be at the individual level. 

```{r, echo=FALSE}
poverty_table %>%
  kbl() %>%
  kable_minimal() %>% 
  kable_styling(html_font = "Unna",
                full_width = FALSE,
                position = "center")


```
```{r, include=FALSE}
#Getting number of HHs in poverty or whatever
#https://cran.r-project.org/web/packages/ipumsr/vignettes/ipums-cps.html
hhs_poverty_status_quo <- ipum %>% 
  group_by(baseline_poverty) %>%
  summarize(hhs_in_poverty = sum(spmwt)) %>% 
  filter(baseline_poverty==TRUE) %>% pull()
hhs_poverty_reform <- ipum %>% 
  group_by(post_reform_poverty) %>%
  summarize(hhs_in_poverty = sum(spmwt)) %>% 
  filter(post_reform_poverty==TRUE) %>% pull()

num_of_hh_out_of_poverty <- hhs_poverty_status_quo-hhs_poverty_reform
num_of_hh_out_of_poverty <- prettyNum(num_of_hh_out_of_poverty,big.mark=",",scientific=FALSE) #Making this display nicely
```
This reform caused a `r num_of_hh_out_of_poverty[1]` household poverty reduction, a `r percent_poverty_reduction[1]`% decrease (a `r percentage_point_poverty_reduction[1]` percentage point reduction). Need to be careful with HH vs. individual level.

## Deep Poverty Impacts of Expanded EITC
```{r, echo=FALSE}
deep_poverty_table %>%
  kbl() %>%
  kable_minimal() %>% 
  kable_styling(html_font = "Unna",
                full_width = FALSE,
                position = "center")
```




```{r, include = FALSE}
#Making sure all the income differences are attributable to the EITC (yes, basically not seeing any dynamic effects)
ipum %>% mutate(eitc_diff=reform_eitc-status_quo_eitc,
                        diff_aftertax_income_no_eitc=eitc_diff-diff_aftertax_income) %>% 
  summarise(across(c(eitc_diff:diff_aftertax_income_no_eitc), mean, na.rm=TRUE)) ## removing nas which maybe should be 0s
```
## Effects Across the Income Distribution

```{r, include = FALSE}
#Need to be careful, this is not a cleaned CPS, for analysis of other variables I need to clean stuff
ipum <- ipum %>% mutate(total_income=ifelse(inctot==999999999,NA,inctot),
                        income_percentile=ntile(total_income,100),
                        income_percentile_20=ntile(total_income,20),
                        income_percentile_20=income_percentile_20*5) 
income_change_percentiles <- ipum %>% group_by(income_percentile_20) %>% 
  summarise(income_change_from_policy=mean(diff_aftertax_income_family))

#Need to figure out how the family unit stuff works
# income_change_percentiles_no_children <- ipum %>% filter() %>% 
#   group_by(income_percentile) %>% 
#   summarise(income_change_from_policy=mean(diff_aftertax_income_family))

income_percentile_change_plot <- ggplot(income_change_percentiles, aes(x = income_percentile_20, y = income_change_from_policy)) +
  geom_line(color = "#572dff") +
  theme_bw() +
  scale_y_continuous(labels=scales::dollar_format()) +
  scale_x_continuous(breaks = seq(0,100,10), labels = seq(0,100,10)) +
    theme(plot.title = element_text(hjust = 0.5),
        text=element_text(family="Unna"), #Want to change this to Jain font, need to ask for it
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(angle=0, vjust = 0.5),
        panel.background = element_rect(fill="#fffdfc"), #Not sure if this is actually modifying background
        plot.background = element_rect(fill="#fffdfc"),
        plot.margin=unit(c(.15,.75,.15,.15),"cm")) +
  labs(y = "Average\nChange\nin\nIncome", x = "Pre-Policy Income Percentile", 
       title = "Average Increase in Income From EITC Reform\nBy Pre-Policy Income Rank",
       caption = "Jain Family Institute")
```


```{r, warning = FALSE, message = FALSE, echo=FALSE, fig.align="center"}
income_percentile_change_plot
```

Should do analysis for different groups, e.g. number of children (that's a HH thing), labor force status, race, more?


```{r, include =FALSE}
test <- ipum %>% filter(labforce!=0 & age>24 & age <65) %>% #prime age labor force or some shit, need to look up the definition
  mutate(labor_force=ifelse(labforce==1,"Not In Labor Force",
                            ifelse(labforce==2,"In Labor Force",NA))) %>% 
  group_by(income_percentile_20,labor_force) %>% #Made this to 20 but should just group 1-5, 6-10 etc.
  summarise(income_change_from_policy=mean(diff_aftertax_income_family, na.rm=TRUE))

#Doesn't really make sense to do this by income percentile for in labor force vs. out of labor force

ggplot(test, aes(x = income_percentile_20, y = income_change_from_policy, group = labor_force)) +
  geom_line(aes(color=as.factor(labor_force))) + #Find complementary colors
  theme_bw() +
  scale_y_continuous(labels=scales::dollar_format()) +
    theme(plot.title = element_text(hjust = 0.5),
        text=element_text(family="Unna"), #Want to change this to Jain font, need to ask for it
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(angle=0, vjust = 0.5),
        panel.background = element_rect(fill="#fffdfc"), #Not sure if this is actually modifying background
        plot.background = element_rect(fill="#fffdfc"),
        plot.margin=unit(c(.15,.75,.15,.15),"cm")) +
  guides(color=guide_legend("")) + 
  labs(y = "Average\nChange\nin\nIncome", x = "Pre-Policy Income Percentile", 
       title = "Average Increase in Income From EITC Reform\nBy Pre-Policy Income Rank And Labor Force Status",
       caption = "Prime Age Workers Only")
```


## Cost of Reform

I think the "total tax revenue" is probably way off due to CPS not being representative of top incomes, so probably should just say the cost rather than including this table with the overall numbers. (The cost difference should be pretty accurate because all the changes are with the low income filers)
```{r, include = FALSE}
#No idea if this is the correct variable/exactly how to do this
cost_table <- ipum %>% summarise(total_tax_rev_status_quo=sum(status_quo_iitax*spmwt, na.rm=TRUE),
                   total_tax_rev_reform=sum(reform_iitax*spmwt, na.rm=TRUE)) %>% 
  mutate(cost_of_reform=total_tax_rev_reform-total_tax_rev_status_quo) %>% 
  mutate(across(where(is.numeric), ~ .x/1000000000)) %>% 
  mutate(across(where(is.numeric), round, 0)) %>% 
  mutate(across(where(is.numeric), ~str_prefix(.x,"$"))) %>% 
  rename(`Status Quo Tax Revenue` = total_tax_rev_status_quo,
         `Total Tax Revenue Reform` = total_tax_rev_reform,
         `Cost of Reform` = cost_of_reform)
```

```{r, echo=FALSE}
cost_table %>%
  kbl() %>%
  kable_minimal() %>% 
  kable_styling(html_font = "Unna",
                full_width = FALSE,
                position = "center") %>% 
  footnote(general = "Cost in Billions of Dollars",
           footnote_as_chunk = TRUE) 
```
<!-- ## Analysis of how non-filing reduces benefits?  -->

<!-- Also need to look at policy with CTC and other stuff recently passed -->


