---
title: "ABC Act Analysis"
author: "Jack Landry"
date: "6/22/2021"
output: html_document
---

<style type="text/css">
body{
  font-family: Unna;
  font-size: 12pt;
}
/* Headers */
h1,h2,h3,h4,h5,h6{
  font-size: 20pt;
  text-align: center;
}

h1.title {
  font-size: 38px;
  text-align: center;
}
h4.author { 
    font-size: 18px;
  text-align: center;
}
h4.date { 
  font-size: 18px;
  text-align: center;
}
</style>

```{r setup, include=FALSE}
#Not sure i
library(tidyverse)
library(ipumsr)
library(tidylog)
library(kableExtra)
library(stringr.tools)
library(here)
here()
#Tax units sent to calculator
ids <- read_csv(here("outputs","ids.csv")) 
#New policy results file
new_policy <- read_csv(here("tax-calculator-outputs","abc-act-reform-taxable.csv"))
#Status quo policy results file
status_quo_poicy <- read_csv(here("tax-calculator-code","c17_taxsim-17-#-#-#.csv"))

#Need to change this to keep all these output variabes, but add a prefix or something
names(new_policy) <- paste("reform", names(new_policy), sep = "_")

# new_policy <- new_policy %>% rename(new_pol_aftertax_income=raftertax_income)
names(status_quo_poicy) <- paste("status_quo", names(status_quo_poicy), sep = "_")
# status_quo_poicy <- status_quo_poicy %>%
#   rename(stats_quo_pol_aftertax_income=aftertax_income)

combo_data <- ids %>% left_join(status_quo_poicy, by = c("RECID" = "status_quo_RECID")) %>% 
  left_join(new_policy, by = c("RECID" = "reform_RECID"))
combo_data <- combo_data %>% mutate(diff_aftertax_income=reform_aftertax_income - status_quo_aftertax_income)

#Now bring in CPS ASEC data raw data for analysis
ddi <- read_ipums_ddi(here("cps_data","cps_00042.xml"))
ipum <-  read_ipums_micro(ddi)
names(ipum) <- tolower(names(ipum))


#"Not every ASEC record became a tax unit, so make sure you keep master non-matches" (From Earnie)
#So ASEC records that are not tax units had no change in income
#Also, why didn't they become tax units???
ipum <- ipum  %>% left_join(combo_data, by = c("serial","pernum"))



```

```{r, include=FALSE}
#Need to adjust everything for inflation somewhere
#Need to note what year of data I'm using and what tax year
#Need to try to do inference somewhere
ipum <- ipum %>% 
  mutate(state_name = labelled::to_factor(ipum$statefip, levels = "labels")) 

ipum <- ipum %>% mutate(no_aftertax_income=ifelse(is.na(diff_aftertax_income),1,0),
                        diff_aftertax_income=ifelse(is.na(diff_aftertax_income),0,diff_aftertax_income)) 

#Adding UBI
ipum <- ipum %>% mutate(diff_aftertax_income=diff_aftertax_income+13000)
#Need to be careful about how missings are treated, I think if they aren't in the tax data the change should be 0
ipum <- ipum %>% group_by(spmfamunit) %>% 
  mutate(diff_aftertax_income_family=sum(diff_aftertax_income)) %>% 
  ungroup()

#SPMTOTRES records the total economic resources of families and is used to construct the Supplemental Poverty Measure (SPM)
#SPMTHRESH records the poverty threshold used to calculate the Supplemental Poverty Measure (SPM) for a family.

#Poverty metrics
ipum <- ipum %>% mutate(deep_poverty=spmthresh*.5,
                baseline_poverty = spmtotres < spmthresh,
                post_reform_poverty = (spmtotres + diff_aftertax_income_family) < spmthresh,
                out_of_poverty = ifelse((post_reform_poverty == 0 & baseline_poverty == 1),1,0),
                out_of_poverty = ifelse(is.na(out_of_poverty),0,out_of_poverty),
                baseline_deep_poverty=spmtotres < deep_poverty,
                post_reform_deep_poverty=(spmtotres + diff_aftertax_income_family) < deep_poverty,
                out_of_deep_poverty = ifelse((post_reform_deep_poverty == 0 & baseline_deep_poverty == 1),1,0),
                out_of_deep_poverty = ifelse(is.na(out_of_deep_poverty),0,out_of_deep_poverty)) 



```

The Automatic BOOST to Communities Act, or the ABC Act, gives a total of \$13,000 to every person in the United States annually until the end of the Coronavirus crisis. ''Every person'' includes children (paid though who claims them as a dependent) and undocumented immigrants. The benefit is destributed monthly, with an initally \$2,000 payment and $1,000 per month thereafter. This document assesses the bill's hypothetical impact on poverty and the income distribution using data from the Current Population Survey's Annual Social and Economic Supplement. 



<!-- ```{r, echo=FALSE, fig.align="center"} -->
<!-- eitc_plot -->
<!-- ``` -->



```{r, include=FALSE}
#Making functions to get poverty by different groups
poverty_by_group_filter <- function(group) {
ipum %>% filter({{ group }}==1) %>% summarise(
  baseline_poverty_mean = weighted.mean(baseline_poverty, spmwt),
  reform_poverty_mean = weighted.mean(post_reform_poverty, spmwt),
  reform_poverty_reduction = weighted.mean(out_of_poverty, spmwt),
  baseline_deep_poverty_mean = weighted.mean(baseline_deep_poverty,spmwt),
  reform_deep_poverty_mean = weighted.mean(post_reform_deep_poverty,spmwt),
  reform_deep_poverty_reduction = weighted.mean(out_of_deep_poverty,spmwt)) 
}
poverty_by_group <- function(group) {
ipum %>% group_by({{ group }}) %>% summarise(
  baseline_poverty_mean = weighted.mean(baseline_poverty, spmwt),
  reform_poverty_mean = weighted.mean(post_reform_poverty, spmwt),
  reform_poverty_reduction = weighted.mean(out_of_poverty, spmwt),
  baseline_deep_poverty_mean = weighted.mean(baseline_deep_poverty,spmwt),
  reform_deep_poverty_mean = weighted.mean(post_reform_deep_poverty,spmwt),
  reform_deep_poverty_reduction = weighted.mean(out_of_deep_poverty,spmwt)) 
}

#Making variables to calculate poverty rates over
ipum <- ipum %>% mutate(child=ifelse(age<18,1,0),
                                     everyone=1)
#Ignoring multi-racial here
ipum <- ipum %>% mutate(race_bin=ifelse(race==100 & hispan==0,"White",
                                ifelse(race==200 & hispan==0,"Black",
                                ifelse(race==651 & hispan==0,"Asian",
                                ifelse(race==100 & hispan!=0,"Hispanic","Other")))))

ipum <- ipum %>% mutate(employment_status=ifelse(empstat<10,NA,
                                         ifelse(empstat>=10 & empstat<=12,"Has Job",
                                                ifelse(empstat>=21 & empstat<=22,"Unenployed",
                                                       ifelse(empstat==32, "Unable to work",
                                                              ifelse(empstat==36, "Retired",NA))))))
#Household size
spm_fam_unit_size_df <- ipum %>% group_by(spmfamunit) %>% count() %>% rename(spm_fam_unit_size=n)

ipum <- ipum %>% left_join(spm_fam_unit_size_df)
ipum %>% group_by(spm_fam_unit_size) %>% count() #Need to cut it off at a certain size
ipum <- ipum %>% mutate(spm_fam_unit_size_cap=ifelse(spm_fam_unit_size>=8,8,spm_fam_unit_size),
                spm_fam_unit_size_cap=as.character(spm_fam_unit_size_cap),
                spm_fam_unit_size_cap=ifelse(spm_fam_unit_size_cap==8,"8+",spm_fam_unit_size_cap))

#Maybe I can do a function for grouping rather than just filtering
child_poverty <- poverty_by_group_filter(child) %>% mutate(Group = "Children")
overall_poverty <- poverty_by_group_filter(everyone) %>% mutate(Group = "Overall")
poverty_by_race <- poverty_by_group(race_bin)
poverty_by_race <- poverty_by_race %>% rename(Group=race_bin)
poverty_by_employment_status <- poverty_by_group(employment_status)
poverty_by_fam_unit_size <- poverty_by_group(spm_fam_unit_size_cap)

#Function to make nice tables
table_maker <- function(dataframe,group_var,group_name) {
dataframe %>% rename(`Status Quo Poverty` = baseline_poverty_mean,
                       `Reform Poverty` = reform_poverty_mean,
                       `Poverty Reduction` = reform_poverty_reduction,
                       `Status Quo Deep Poverty` = baseline_deep_poverty_mean,
                       `Reform Deep Poverty` = reform_deep_poverty_mean,
                       `Deep Poverty Reduction` = reform_deep_poverty_reduction,
                       {{group_name}} := {{group_var}}) %>% 
  mutate(across(where(is.numeric), round, 3)) %>% 
  mutate(across(where(is.numeric), ~ .x*100)) %>% 
  mutate(across(where(is.numeric),as.character)) %>% 
  mutate(across(`Status Quo Poverty`:`Deep Poverty Reduction`, ~str_suffix(.x,"%"))) %>% 
  relocate({{group_name}})

}
#Applying function and a bit of manual cleanu
poverty_by_fam_unit_size <- table_maker(poverty_by_fam_unit_size,spm_fam_unit_size_cap,`Family Size`)

combo_poverty <- bind_rows(overall_poverty,child_poverty,poverty_by_race) 
combo_poverty <- table_maker(combo_poverty,Group,Group)
poverty_by_employment_status <- table_maker(poverty_by_employment_status,employment_status,`Employment Status`)
poverty_by_employment_status <- poverty_by_employment_status %>% filter(!is.na(`Employment Status`))
#Lets do overall, children, and race in one table, employment status in another table, and fam unit size in another?
#Maybe I should make some graphs here... Like maybe arrows showing poverty reductions


#Might be interesting to show how this can mute poverty impacts in a recession, e.g look over time
#Who are the remaining poor people... mostly negative total resource measure under status quo, also need to look more into how the threshold is calculated
# ipum %>% filter(post_reform_poverty==1) %>% select(inctot,diff_aftertax_income,diff_aftertax_income_family,spmthresh,spmtotres)

#Need to make sure this is still correct
percent_poverty_reduction <- (overall_poverty$baseline_poverty_mean-overall_poverty$reform_poverty_mean)/overall_poverty$baseline_poverty_mean

percent_poverty_reduction <- round(percent_poverty_reduction, 3)*100
percentage_point_poverty_reduction <- round(overall_poverty$reform_poverty_reduction, 3)*100
```
<!-- Calculating changes in poverty, want to make these nice tables -->
<!-- Want to change the html to mimic Jain style as much as possible -->
<!-- Look at example here: https://www.jainfamilyinstitute.org/news/robust-evidence-for-1400-relief-and-recovery-checks/ -->



## Poverty Impacts of the ABC Act By Group


```{r, include=FALSE}
#Exploring the people who are still poor after the refuorm
post_reform_poor <- ipum %>% filter(post_reform_poverty==1)
ggplot(post_reform_poor, aes(x = spmthresh)) +
  geom_histogram()
#Unclear what's going on here
ggplot(post_reform_poor, aes(x = status_quo_aftertax_income)) +
  geom_histogram()
```



```{r, echo=FALSE}
combo_poverty %>%
  kbl() %>%
  kable_minimal() %>% 
  kable_styling(html_font = "Unna",
                full_width = FALSE,
                position = "center")
```

```{r, echo=FALSE}
poverty_by_employment_status %>% 
  kbl() %>% 
  kable_minimal() %>% 
  kable_styling(html_font = "Unna",
                full_width = FALSE,
                position = "center")
```

```{r, echo=FALSE}
poverty_by_fam_unit_size %>% 
  kbl() %>% 
  kable_minimal() %>% 
  kable_styling(html_font = "Unna",
                full_width = FALSE,
                position = "center")


```

```{r, include=FALSE}
#Getting number of HHs in poverty or whatever
#https://cran.r-project.org/web/packages/ipumsr/vignettes/ipums-cps.html
hhs_poverty_status_quo <- ipum %>% 
  group_by(baseline_poverty) %>%
  summarize(hhs_in_poverty = sum(spmwt)) %>% 
  filter(baseline_poverty==TRUE) %>% pull()
hhs_poverty_reform <- ipum %>% 
  group_by(post_reform_poverty) %>%
  summarize(hhs_in_poverty = sum(spmwt)) %>% 
  filter(post_reform_poverty==TRUE) %>% pull()

num_of_hh_out_of_poverty <- hhs_poverty_status_quo-hhs_poverty_reform
num_of_hh_out_of_poverty <- prettyNum(num_of_hh_out_of_poverty,big.mark=",",scientific=FALSE) #Making this display nicely
```
Overall, the ABC Act would cause a `r num_of_hh_out_of_poverty[1]` household poverty reduction, a `r percent_poverty_reduction[1]`% decrease (a `r percentage_point_poverty_reduction[1]` percentage point reduction). 



## ABC Act's Distributional Impact

The ABC ACT disproportunatly increases the income of low income families, even though there is no means testing or phase out. The mechanism for this effect is family size. Low income families are on average larger, so they see a larger benefit on average.

```{r, include = FALSE}
#Need to be careful, this is not a cleaned CPS, for analysis of other variables I need to clean stuff
ipum <- ipum %>% mutate(total_income=ifelse(inctot==999999999,NA,inctot),
                        income_percentile=ntile(total_income,100),
                        income_percentile_20=ntile(total_income,20),
                        income_percentile_20=income_percentile_20*5) 
#Why are these numbers so big
ipum %>% select(diff_aftertax_income_family,income_percentile_20,spm_fam_unit_size) #Why is income percentile NA sometimes
#How about adding some cofidence bands to this sucker
#Could use the survey package, maybe better myself?
income_change_percentiles <- ipum %>% group_by(income_percentile_20) %>% 
  summarise(income_change_from_policy=mean(diff_aftertax_income_family), #This should be weighted mean anyway
            spm_family_size_mean=mean(spm_fam_unit_size))
```

```{r, include = FALSE}
library(srvyr)
ipum_spmwt <- ipum %>%
   as_survey_design(weights = spmwt)
income_change_percentiles_ci <- ipum_spmwt %>% group_by(income_percentile_20) %>% 
  summarise(income_change_from_policy=survey_mean(diff_aftertax_income_family, vartype = "ci"), #This should be weighted mean anyway
            spm_family_size_mean=survey_mean(spm_fam_unit_size, vartype = "ci"))


#Need to figure out how the family unit stuff works
# income_change_percentiles_no_children <- ipum %>% filter() %>% 
#   group_by(income_percentile) %>% 
#   summarise(income_change_from_policy=mean(diff_aftertax_income_family))
```



```{r, include = FALSE}
income_percentile_change_plot <- ggplot(income_change_percentiles_ci, aes(x = income_percentile_20, y = income_change_from_policy)) +
  geom_line(color = "#572dff") +
  geom_path(color = "#572dff") +
  geom_ribbon(aes(ymin=income_change_from_policy_low, ymax=income_change_from_policy_upp), linetype=2, alpha=0.2) +
  theme_bw() +
  scale_y_continuous(labels=scales::dollar_format(), breaks = seq(0,60000,15000), limits = c(0, 60000)) + #Breaks and labels with scales oh my
  scale_x_continuous(breaks = seq(0,100,10), labels = seq(0,100,10)) +
    theme(plot.title = element_text(hjust = 0.5),
        text=element_text(family="Unna"), #Want to change this to Jain font, need to ask for it
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(angle=0, vjust = 0.5),
        panel.background = element_rect(fill="#fffdfc"), #Not sure if this is actually modifying background
        plot.background = element_rect(fill="#fffdfc"),
        plot.margin=unit(c(.15,.75,.15,.15),"cm")) +
  labs(y = "Average\nChange\nin\nIncome", x = "Pre-Policy Income Percentile", 
       title = "Average Increase in Income From ABC Act\nBy Pre-Policy Income Rank",
       caption = "Jain Family Institute")

family_size_by_income <- ggplot(income_change_percentiles_ci, aes(x = income_percentile_20, y = spm_family_size_mean)) +
  geom_line(color = "#572dff") +
    geom_path(color = "#572dff") +
  geom_ribbon(aes(ymin=spm_family_size_mean_low, ymax=spm_family_size_mean_upp), linetype=2, alpha=0.2) +
  theme_bw() +
  scale_y_continuous(labels = seq(0,5,1), breaks = seq(0,5,1), limits = c(0,5)) +
  scale_x_continuous(breaks = seq(0,100,10), labels = seq(0,100,10)) +
    theme(plot.title = element_text(hjust = 0.5),
        text=element_text(family="Unna"), #Want to change this to Jain font, need to ask for it
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(angle=0, vjust = 0.5),
        panel.background = element_rect(fill="#fffdfc"), #Not sure if this is actually modifying background
        plot.background = element_rect(fill="#fffdfc"),
        plot.margin=unit(c(.15,.75,.15,.15),"cm")) +
  labs(y = "Average\nFamily\nSize", x = "Pre-Policy Income Percentile", 
       title = "Average Family Size \nBy Pre-Policy Income Rank",
       caption = "Jain Family Institute")

```


```{r, warning = FALSE, message = FALSE, echo=FALSE, fig.align="center"}
income_percentile_change_plot
```

```{r, warning = FALSE, message = FALSE, echo=FALSE, fig.align="center"}
family_size_by_income
```



```{r, include =FALSE}
test <- ipum %>% filter(labforce!=0 & age>24 & age <65) %>% #prime age labor force or some shit, need to look up the definition
  mutate(labor_force=ifelse(labforce==1,"Not In Labor Force",
                            ifelse(labforce==2,"In Labor Force",NA))) %>% 
  group_by(income_percentile_20,labor_force) %>% #Made this to 20 but should just group 1-5, 6-10 etc.
  summarise(income_change_from_policy=mean(diff_aftertax_income_family, na.rm=TRUE))

#Doesn't really make sense to do this by income percentile for in labor force vs. out of labor force

ggplot(test, aes(x = income_percentile_20, y = income_change_from_policy, group = labor_force)) +
  geom_line(aes(color=as.factor(labor_force))) + #Find complementary colors
  theme_bw() +
  scale_y_continuous(labels=scales::dollar_format()) +
    theme(plot.title = element_text(hjust = 0.5),
        text=element_text(family="Unna"), #Want to change this to Jain font, need to ask for it
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(angle=0, vjust = 0.5),
        panel.background = element_rect(fill="#fffdfc"), #Not sure if this is actually modifying background
        plot.background = element_rect(fill="#fffdfc"),
        plot.margin=unit(c(.15,.75,.15,.15),"cm")) +
  guides(color=guide_legend("")) + 
  labs(y = "Average\nChange\nin\nIncome", x = "Pre-Policy Income Percentile", 
       title = "Average Increase in Income From EITC Reform\nBy Pre-Policy Income Rank And Labor Force Status",
       caption = "Prime Age Workers Only")
```


## Cost of Reform

<!-- I think the "total tax revenue" is probably way off due to CPS not being representative of top incomes, so probably should just say the cost rather than including this table with the overall numbers. (The cost difference should be pretty accurate because all the changes are with the low income filers) -->


```{r, include = FALSE}
#No idea if this is the correct variable/exactly how to do this
total_cost_trillions <- ipum %>% summarise(total_cost=sum(diff_aftertax_income*spmwt, na.rm=TRUE)) %>% 
  mutate(total_cost_trillions = total_cost/1000000000000,
         total_cost_trillions = round(total_cost_trillions,4)) %>% 
  select(total_cost_trillions)
total_cost_trillions
# cost_table <- ipum %>% summarise(total_tax_rev_status_quo=sum(status_quo_iitax*spmwt, na.rm=TRUE),
#                    total_tax_rev_reform=sum(reform_iitax*spmwt, na.rm=TRUE)) %>% 
#   mutate(cost_of_reform=total_tax_rev_reform-total_tax_rev_status_quo) %>% 
#   mutate(across(where(is.numeric), ~ .x/1000000000)) %>% 
#   mutate(across(where(is.numeric), round, 0)) %>% 
#   mutate(across(where(is.numeric), ~str_prefix(.x,"$"))) %>% 
#   rename(`Status Quo Tax Revenue` = total_tax_rev_status_quo,
#          `Total Tax Revenue Reform` = total_tax_rev_reform,
#          `Cost of Reform` = cost_of_reform)
```

Annually, the ABC act would cost approximately $`r total_cost_trillions[1]` trillion dollars. This assumes none of the benefit is taxable, and no one changes their behavior in response to the benefit (e.g. reducing labor supply).




<!-- ```{r, echo=FALSE} -->
<!-- cost_table %>% -->
<!--   kbl() %>% -->
<!--   kable_minimal() %>%  -->
<!--   kable_styling(html_font = "Unna", -->
<!--                 full_width = FALSE, -->
<!--                 position = "center") %>%  -->
<!--   footnote(general = "Cost in Billions of Dollars", -->
<!--            footnote_as_chunk = TRUE)  -->
<!-- ``` -->
<!-- <!-- ## Analysis of how non-filing reduces benefits?  -->  

<!-- Also need to look at policy with CTC and other stuff recently passed -->


